Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-03-22 manuscript/supplemental_material.Rnw
use FPKM=1 for most gene trees; add PE genes

\begin{figure}[hb]
<<adam12.evolution.tree,echo=FALSE,out.height="0.6\\textheight",fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="ADAM12" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=5,
                 axis.text.size=5,
                 fontsize=8
                 )
@ 
\caption{\genetreeexplanation{\textit{ADAM12}}}
\label{fig:adam12_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="KISS1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=5,
                 axis.text.size=5,
                 axis.text.angle=0,
                 fontsize=5
                 )
@ 
\caption{\genetreeexplanation{\textit{KISS1}}}
\label{fig:kiss1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PAPPA" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=5,
                 axis.text.size=5,
                 axis.text.angle=0,
                 fontsize=5
                 )
@ 
\caption{\genetreeexplanation{\textit{PAPPA}}}
\label{fig:pappa_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="TGFB1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 subtree=c("ENSMODG00000008630","ENSSSCG00000026296"),
                 gene.tree.table,gene.tree.expression,min.fpkm=5,
                 axis.text.size=5,
                 fontsize=5
                 )
@ 
\caption{\genetreeexplanation{\textit{TGFB1}}}
\label{fig:tgfb1_evolution_tree}
\end{figure}

\begin{figure}
<<gpc3.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="GPC3" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=5,
                 axis.text.size=5,
                 fontsize=8
                 )
@ 
\caption{\genetreeexplanation{\textit{GPC3}}}
\label{fig:gpc3_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="FSTL1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=5,
                 axis.text.size=5,
                 fontsize=8
                 )
@ 
\caption{\genetreeexplanation{follistatin-like 1 \textit{FSTL1}}}
\label{fig:fstl1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="SERPINE2" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=1,
                 axis.text.size=5,
                 fontsize=4)
@ 
\caption{\genetreeexplanation{\textit{SERPINE2}}}
\label{fig:serpine_evolution_tree}
\end{figure}

>>>>

\begin{figure}[hb]
<<crh.evolution.tree,echo=FALSE,out.height="0.6\\textheight",fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CRH" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 min.fpkm=1,
                 axis.text.size=5,
                 axis.text.angle=0,
                 fontsize=8
                 )
@ 
\caption{\genetreeexplanation{\textit{CRH}}}
\label{fig:crh_evolution_tree}
\end{figure}

\begin{figure}[hb]
<<adam12.evolution.tree,echo=FALSE,out.height="0.6\\textheight",fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="ADAM12" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,
                 gene.tree.expression,
                 subtree=c("ENSMODG00000010538","ENSSSCG00000010746"),
                 min.fpkm=0.1,
                 axis.text.size=4,
                 fontsize=5
                 )
@ 
\caption{\genetreeexplanation{\textit{ADAM12}}}
\label{fig:adam12_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="KISS1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 min.fpkm=0.1,
                 axis.text.size=5,
                 axis.text.angle=0,
                 fontsize=5
                 )
@ 
\caption{\genetreeexplanation[0.1]{\textit{KISS1}}}
\label{fig:kiss1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PAPPA" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 subtree=c("ENSSSCG00000005498","ENSG00000182752"),
                 min.fpkm=0.1,
                 axis.text.size=5,
                 axis.text.angle=0,
                 fontsize=5
                 )
@ 
\caption{\genetreeexplanation[0.1]{\textit{PAPPA}}}
\label{fig:pappa_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="TGFB1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 subtree=c("ENSMODG00000008630","ENSSSCG00000026296"),
                 gene.tree.table,gene.tree.expression,
                 min.fpkm=1,
                 axis.text.size=3.5,
                 fontsize=5
                 )
@ 
\caption{\genetreeexplanation{\textit{TGFB1}}}
\label{fig:tgfb1_evolution_tree}
\end{figure}

\begin{figure}
<<gpc3.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="GPC3" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,
                 subtree=c("ENSCAFG00000018864","ENSLAFG00000025706"),
                 gene.tree.expression,
                 min.fpkm=1,
                 axis.text.size=3.5,
                 fontsize=6
                 )
@ 
\caption{\genetreeexplanation{\textit{GPC3}}}
\label{fig:gpc3_evolution_tree}
\end{figure}

\subsubsection{Primate Specific Evolution}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="FSTL1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,
                 gene.tree.expression,
                 min.fpkm=1,
                 axis.text.size=5,
                 fontsize=8
                 )
@ 
\caption{\genetreeexplanation{follistatin-like 1 \textit{FSTL1}}}
\label{fig:fstl1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="SERPINE2" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,
                 gene.tree.expression,
                 min.fpkm=1,
                 axis.text.size=5,
                 fontsize=4)
@ 
\caption{\genetreeexplanation{\textit{SERPINE2}}}
\label{fig:serpine_evolution_tree}
\end{figure}
